The Emergent Simplicity of Galaxy Size

### 2017-09-15 paper/galsize_paper.tex
Finished description of size modeling

%%%%%%%%  galaxy properties  %%%%%%%%
\newcommand{\rhalf}{R_{1/2}}
\newcommand{\rhalfdisk}{R_{1/2}^{\rm disk}}
\newcommand{\rhalfbulge}{R_{1/2}^{\rm bulge}}
\newcommand{\adisk}{A_{\rm disk}}
\newcommand{\abulge}{A_{\rm bulge}}
\newcommand{\alphadisk}{\alpha_{\rm disk}}
\newcommand{\alphabulge}{\alpha_{\rm bulge}}
\newcommand{\sigmarhalf}{\sigma_{\rm R_{1/2}}}
\newcommand{\rvir}{R_{\rm vir}}
\newcommand{\bt}{{\rm B/T}}
\newcommand{\mstar}{M_{\ast}}
\newcommand{\ssfr}{{\rm sSFR}}
\newcommand{\sfr}{{\rm SFR}}

%%%%%%%%  halo properties  %%%%%%%%
\newcommand{\halospin}{\lambda_{\rm halo}}
\newcommand{\mvir}{M_{\rm vir}}
\newcommand{\macc}{M_{\rm acc}}
\newcommand{\mpeak}{M_{\rm peak}}
\newcommand{\mhalo}{M_{\rm halo}}

@@

Motivated by the \citet{kravtsov13} results, we model the half-mass radii of galaxies as separate power law functions of halo virial radius:

\begin{eqnarray}
\label{eq:fiducial_model}
\rhalfbulge &=& \abulge\rvir^{\alphabulge} \\
\rhalfdisk &=& \adisk\rvir^{\alphadisk}
\end{eqnarray}

For central galaxies residing in host halos, we use the present-day virial radius; for satellite galaxies we use the virial radius at the time of accretion, using physical units of $\kpc.$ We model the composite size of model galaxies as
\beq
\rhalf = (\bt)\rhalfbulge + (1-\bt)\rhalfdisk.
\eeq

When generating Monte Carlo realizations of our model galaxy population, we add log-normal scatter $\sigmarhalf$ separately to each component. Our model for sizes thus has five free parameters: $\abulge, \adisk, \alphabulge, \alphadisk,$ and $\sigmarhalf.$

We reiterate an important distinction drawn in the formulation of our forward modeling approach. In the model described above, $\rhalfbulge$ and $\rhalfdisk$ are forward modeling quantities that we aim to constrain through SDSS observations of {\em composite} galaxy size $\rhalf.$ In particular, both model and real galaxies will be categorized as ``small" or ``large`` according to whether $\rhalf$ is above or below the median size for a galaxy with total stellar mass $\mstar.$

>>>>

%%%%%%%%  galaxy properties  %%%%%%%%
\newcommand{\rhalf}{R_{1/2}}
\newcommand{\rhalfdisk}{R_{1/2}^{\rm disk}}
\newcommand{\rhalfbulge}{R_{1/2}^{\rm bulge}}
\newcommand{\adisk}{A_{\rm disk}}
\newcommand{\abulge}{A_{\rm bulge}}
\newcommand{\alphadisk}{\alpha_{\rm disk}}
\newcommand{\alphabulge}{\alpha_{\rm bulge}}
\newcommand{\sigmarhalf}{\sigma_{\rm R_{1/2}}}
\newcommand{\bt}{{\rm B/T}}
\newcommand{\mstar}{M_{\ast}}
\newcommand{\ssfr}{{\rm sSFR}}
\newcommand{\sfr}{{\rm SFR}}

%%%%%%%%  halo properties  %%%%%%%%
\newcommand{\halospin}{\lambda_{\rm halo}}
\newcommand{\mvir}{M_{\rm vir}}
\newcommand{\macc}{M_{\rm acc}}
\newcommand{\mpeak}{M_{\rm peak}}
\newcommand{\mhalo}{M_{\rm halo}}
\newcommand{\rvir}{R_{\rm vir}}
\newcommand{\rmpeak}{R_{\rm M_{peak}}}

@@

In \citet{kravtsov13}, it was found that if a stellar-to-halo mass relation is inverted to map halo mass estimates $\mhalo$ onto SDSS galaxies, and then the $\mhalo-\rvir$ relation is applied to map values of $\rvir$ onto the galaxies, then the resulting $\rhalf-\rvir$ relation of SDSS galaxies exhibits the following linear scaling across a wide range of stellar mass:
\beq
\label{eq:fiducial_model}
\rhalf = 0.0125\rvir
\eeq

Motivated by the simplicity of this scaling relation, we transform the \citet{kravtsov13} into a forward model using {\tt Halotools}. For the virial radius of halos and subhalos, we use $\rmpeak,$ the value of $\rvir$ in physical units of $\kpc$ measured at the time of peak subhalo mass. When generating Monte Carlo realizations of our model galaxy sizes, we add log-normal scatter $\sigmarhalf=0.15$ dex.
